---
title: "20250403_ihs_SW_v3"
author: "Johanne"
date: "2025-04-03"
output: html_document
---
```{r loading packages, include=FALSE, suppressPackageStartupMessages = FALSE}
library(ggplot2)
library(tidyr)
library(effsize)
library(ggvenn)
library(birk)
library(sjmisc)
library(gtools)
```

This script is what has been used to create the sliding windows for iHS used in 
the paper. 


# Loading raw data 

## General info 
```{r loading info, include = FALSE}
#loading population names
pops <- read.table('/path/to/prefix_pops.txt')
pops <- pops$V1
pops <- pops[which(pops!="LPR")] #removing LPR from analysis

#array names 
ALTAI <- pops[1:12]
SOGEN1 <- pops[13:21]
SOGEN25 <- pops[22:25]

#loading bin info
bins <- read.table('/path/to/bin/number_ind.txt', header=FALSE)
bins <- bins[c(1:21, 23:26),] #removing LPR from analysis
bins <- as.vector(bins)
```

Here, the raw ihh values computed by selscan are transformed into absolute values, 
the pvalue is computed according to Gautier et al., 2017. 


```{r loading files}
to_remove <- NULL  #will be used to clean up
raw_data <- list()

for(p in pops){
  assign(paste0(p), data.frame())
  for(k in seq(1:22)){
    #loading data
    assign(paste0('ihh_LD_' , k), read.table(paste0('/path/to/selscan/output/', p, '/DATE_selscan.', p ,'.', k, '.ihs.out.', bins[match(p, pops)], 'bins.norm' ), col.names = c("locusID", "POSITION", "freq", "sl1", "sl2", 'unstandardized_iHS', 'iHS', 'zscore')))
    
    
    abs_ihh <- abs(get(paste0('ihh_LD_' , k))$iHS) #absolute values of iHH values
    chrom <- rep(k, times=dim(get(paste0('ihh_LD_' , k)))[1])     #add chromosome
    pval <- -log10(2*pnorm(-abs(get(paste0('ihh_LD_' , k))$iHS))) #compute p-value
    
    assign(paste0('ihh_LD_' , k), cbind(chrom, get(paste0('ihh_LD_' , k)), abs_ihh, pval))
    to_remove <- c(to_remove, paste0('ihh_LD_',k))
    assign(paste0(p), rbind(get(paste0(p)), get(paste0('ihh_LD_' , k))))
  }
}
```

```{r cleanup 1, include=F}
raw_data <- list(AKI,AKZ,BOU,HKS,KAZ,KIB,SHO,TAB,TLG,TUB,TUR,UZB,CBR,CJO,CKC,CMR,CPN,CRE,CSA,CST,CTA,LKM,LMT,LPO,LTA)
names(raw_data) <- pops
rm(AKI,AKZ,BOU,HKS,KAZ,KIB,SHO,TAB,TLG,TUB,TUR,UZB,CBR,CJO,CKC,CMR,CPN,CRE,CSA,CST,CTA,LKM,LMT,LPO,LTA)
to_remove <- unique(to_remove)
rm(list=to_remove)
rm(pval, to_remove)
```

# Sliding windows 
## Extremities 
The goal here is to find the furthest start and end for each pop. This will be 
used to start the first window and stop the last. 
```{r closest and furthest position all array}
##Generalizing sliding windows for all arrays 
begining <- data.frame()
ending <- data.frame()
for(p in pops){
  df <- raw_data[[p]]
  for(k in seq(1,22)){
    sub.df <- df[which(df$chrom == k),]
    begining[p,k] <- sub.df$POSITION[which.min(sub.df$POSITION)] #first position
    ending[p,k] <-  sub.df$POSITION[which.max(sub.df$POSITION)] #last position
  }
}
first <- apply(X=begining, MARGIN = 2, min) #closest position
last <- apply(X=ending, MARGIN = 2, max) #furthest position
rm(begining, ending, df, sub.df) #cleanup
```

## Slicing
Slicing into 100kb windows
```{r dividing in intervals}
names.100 <- c()

#p counts the pop names 
for(p in pops){
  print(paste("Processing population : ", p))
  df <- raw_data[[p]]
  quantile_nsl <- quantile(df$abs_ihh, 0.99) #top 1% threshold genome-wide
  assign(paste0(p, '_SW.100'), data.frame())
  
  #k counts the chromosomes 
  for(k in seq(1:22)){
    print(paste("Chromosome treated :",k))
    sub.df <- df[which(df$chrom == k),]
    sub.df <- sub.df[order(sub.df$chrom, sub.df$POSITION),] #ordering
    
    ###100kb###
    w_beg <- first[k] #begining of 100kb interval
    w_end <- w_beg + 100000 #end of 100kb interval
    while(w_end < last[k]){
      nb_snp <- as.numeric(length(which(sub.df$POSITION > w_beg & sub.df$POSITION < w_end))) #number of SNPs in 100kb
      top_snp <- as.numeric(length(which(sub.df$POSITION > w_beg & sub.df$POSITION < w_end & sub.df$abs_ihh >= quantile_nsl))) #nb of snps in top 1%
      prop_snp <- top_snp/nb_snp
      assign(paste0(p, '_SW.100'), rbind(get(paste0(p, '_SW.100')), c(k, w_beg, w_end, nb_snp, top_snp, prop_snp))) #final table for 100kb
      w_beg <- w_beg + 50000
      w_end <- w_beg + 100000
      names.100 <- c(names.100, paste0(p, '_SW.100'))
    }
    #print(paste("End of 100kb sectionning"))
  }
}

col_names <- c("chrom", "w_beg", "w_end", "nb_snp", "top_snp", "prop_snp")

for(p in pops) {
  colnames(SW.100[[p]]) <- col_names
}
```    


## Outliers
Here we harvest the windows that show a proportion of SNPs within the top 1%
distribution. We also filter and remove windows that contain 5 SNPs or less. 

```{r window outliers}
names.100 <- c()

#p counts the pop names 
for(p in pops){
  
  #100kb
  threshold <- 5 #if 5 SNPs or less, window is discared. 
  assign(paste0(p, '.outlier.100'), data.frame())
  
  df <- SW.100[[p]]
  df <- df[-which(df$nb_snp <= threshold),]
  quantile.win <- quantile(df$prop_snp, 0.99, na.rm =TRUE) #top 1% of prop_snp
  assign(paste0(p, '.outlier.100'), df[which(df$prop_snp >= quantile.win),])
  names.100 <- c(names.100, paste0(p, '.outlier.100'))
}

```

Now we have a file of outlier windows for each population. 


# MOST SHARED WINDOWS
We decided to find the outliers that were most shared between populations. 

## Preparation: aggregation of all outlier data 
To do so, we used bedtools to intersect all the outlier window coordinates. 

```{r exporting results}
#p counts the pop names 
for(p in pops){
  
  #100kb
  df <- outliers.100[[p]]
  write.table(x = df[,1:3], paste0("/path/to/output/DATE_outliers_100kb.", p, ".bed"), sep = '\t', quote = FALSE, row.names = FALSE, col.names = FALSE)
}
```

```{bash}
source /path/to/conda.sh
conda activate bedtools

#100kb
bedtools multiinter -i /path/to/output/DATE_outliers_100kb.*.bed -names AKI AKZ BOU CBR CJO CKC CMR CPN CRE CSA CST CTA HKS KAZ KIB LKM LMT LPO LTA SHO TAB TLG TUB TUR UZB -header > /path/to/intersect.txt
```

This process results in a file compiling all intersections between all of the 25 
sets of outlier windows. It gets imported back into R. 

```{r importing results}
directory <- 'path/to/intersect.txt'
intersect.100 <- read.table(directory, header=TRUE)

#keep lines shared by at least 2 populations. 
intersect.100 <- intersect.100[-which(intersect.100$num == 1),]
```


## Concatenation

We decided to concatenate to reduce the bulk of the data. Genomic coordinates of
the windows get concatenated if the windows 1) overleap or 2) touch (have the 
same bounds). 

```{r concatenation}
#init
win.concat.100 <- data.frame()
df <- intersect.100
table_end <- dim(df)[1]-1
final_beg = df$start[1]
final_end = df$end[1]
pops_included <- c()
freq_pops <- c()

#i iterates through the lines of intesect.txt
for(i in c(1:table_end)){
  if (df$chrom[i] == df$chrom[i+1]){ #if we stay on the same chrom
    if(df$start[i+1] <= df$end[i]){ #if end of windows within start of the next
      final_end = df$end[i+1] #extend window
      freq_pops <- c(freq_pops, df$num[i])
      pops_included <- c(pops_included, unlist(strsplit(intersect.100$list[i], split=',')))
      #print(paste( "touching : " , " beg = " , final_beg, "end = ", final_end))
    } else {
      freq_pops <- c(freq_pops, df$num[i])
      pops_included <- c(pops_included, unlist(strsplit(intersect.100$list[i], split=',')))
      #print(paste(unique(pops_included), collapse = " "))
      #print(paste("not touching : ", "beg = ", final_beg, "end = ", final_end))
      win.concat.100 <- rbind(win.concat.100, c(df$chrom[i], final_beg, final_end, max(freq_pops), c(paste(unique(pops_included), collapse = " "))))
      freq_pops <- NULL
      pops_included <- NULL 
      final_beg <- df$start[i+1]
      final_end <- df$end[i+1]
    }
  } else {
    freq_pops <- c(freq_pops, df$num[i])
    pops_included <- c(pops_included, unlist(strsplit(intersect.100$list[i], split=',')))
    #print(pops_included)
    #print(paste(unique(pops_included), collapse = " "))
    #print(paste("changing chrom : ", final_beg, final_end))
    win.concat.100 <- rbind(win.concat.100, c(df$chrom[i], final_beg, final_end, max(freq_pops), c(paste(unique(pops_included), collapse = " "))))
    freq_pops <- NULL
    pops_included <- NULL
    final_beg <- df$start[i+1]
    final_end <- df$end[i+1]
  }
}
names(win.concat.100) <- c('chrom', 'w_beg', 'w_end', 'max_freq', 'list')
head(win.concat.100)
rm(df, table_end, final_beg, final_end, pops_included, freq_pops)
```

Because the vector adding all the pop names was hard to count, we add back 
the actual number of populations per line. 

```{r add number of shared pops}
freq <- c()
#w counts windows 
for(w in c(1:dim(win.concat.100)[1])){
  freq <- c(freq, as.numeric(length(unlist(strsplit(win.concat.100$list[w], split = ' ')))))
}
win.concat.100 <- cbind(win.concat.100, freq)
```


## Prop_snp recalculation

The windows have new bounds now. Therefore the prop_snp calculation is to be 
redone. This is due to the 50% redundancy between windows: half of the SNPs used 
to compute prop_snp in ex-window 1 is also used in ex-window 2 therefore 
new-window 12 does not equate to prop_snp1 + prop_snp 2. 


```{r recalculating 100kb}
#init
prop.snp.100 <- data.frame()
nb.snp.100 <- data.frame()
top.snp.100 <- data.frame()
prop_snp <- c()
nb_snp <- c()
top_snp <- c()
quantile.pops.100 <- c()

#w counts windows 
for(w in c(1:dim(win.concat.100)[1])){
  w_beg <- as.numeric(win.concat.100[w,2])  #init
  w_end <- as.numeric(win.concat.100[w,3])  #init
  chrom <- as.numeric(win.concat.100[w,1])  #init
  
  #p counts pop names 
  for(p in pops) {
    df <- raw_data[paste(p)][[1]]
    quantile_ihs <- quantile(df$abs_ihh, 0.99) #top 1%
    region <- df[which(df[,3] >= w_beg & df[,3] <= w_end & df[,1] == chrom),] #subset to region
    
    nb_snp[p] <- as.numeric(dim(region)[1]) 
    top_snp[p] <- as.numeric(length(which(region$abs_ihh >= quantile_ihs)))
    prop_snp[p] <- top_snp[p]/nb_snp[p]
    quantile.pops.100[p] <- quantile_ihs
  }
  prop.snp.100 <- rbind(prop.snp.100, c(chrom, w_beg, w_end, prop_snp, win.concat.100$list[w]))
  nb.snp.100 <- rbind(nb.snp.100, c(chrom, w_beg, w_end, nb_snp, win.concat.100$list[w]))
  top.snp.100 <- rbind(top.snp.100, c(chrom, w_beg, w_end, top_snp, win.concat.100$list[w]))
}
names(prop.snp.100) <- c(names(win.concat.100[1:3]), pops, "list")
names(nb.snp.100) <- c(names(win.concat.100[1:3]), pops, "list")
names(top.snp.100) <- c(names(win.concat.100[1:3]), pops, "list")
```

## Most shared pops
```{r most shared }
most.shared <- win.concat.100[which(win.concat.100$freq >= 16),]
```



# ECO REGIONAL TEST 
Here we want to identify if there are any signals shared by all pops in NCA but 
none SEA or vice versa. We want to test that using a statistical test. 

## Non parametric tests

Here we use a file containing latitudinal and subsistence strategyinfo for all 
pops, that we then divide into North and South categories for Ecoregional testing. 

```{r loading pop info, include=FALSE}
directory <- '/path/to/pop/context_pops.txt'

pop.info <- read.table(directory, sep = '\t', header=TRUE)
lat_cat <- rep('NA', times = dim(pop.info)[1])
lat_cat[which(pop.info$Lat > 35)] <- 'North'
lat_cat[which(pop.info$Lat < 35)] <- 'South'
pop.info <- cbind(pop.info, lat_cat)
```

Here we test to see whether the proportion of SNPs is significantly different 
between North and South groups. 

```{r NP for latitude, warning=FALSE}
#init
df <- prop.snp.100
lat.100 <- data.frame()

#w counts the windows 
for(w in c(1:dim(df)[1]))
{
  prop_snp <- c(unlist(df[w,4:28]))
  prop_snp <- as.numeric(prop_snp)
  
  data_lm <- cbind(prop_snp, pop.info)
  data_lm$lat_cat <- as.factor(data_lm$lat_cat)
  
  #Condition : more than 2 values are different from 0
  if(as.numeric(length(which(data_lm$prop_snp != 0))) >= 2 & table(is.na(data_lm$prop_snp))[[1]] == 25)
  {
    test <- wilcox.test(prop_snp~lat_cat, data_lm)
    #compute pvalue for latitude
    pval.lat <- test$p.value
    # Extract the W statistic
    W <- test$statistic
    # Sizes of the groups
    n1 <- sum(data_lm$lat_cat == "North")
    n2 <- sum(data_lm$lat_cat == "South")
    # Calculate U1 and U2
    U1 <- W - (n1 * (n1 + 1)) / 2
    U2 <- n1 * n2 - U1
    # Calculate the U statistic (the smaller of U1 and U2)
    U <- min(U1, U2)
    size.effect <- cliff.delta(prop_snp ~ lat_cat, data = data_lm)$estimate
  } 
  else 
  {
    pval.lat <- NA
    W <- NA
    U <- NA
    size.effect <- NA
  }
  lat.100 <- rbind(lat.100, c(df[w,1:3], pval.lat, W, U, size.effect))
  names(lat.100) <- NULL
}
names(lat.100) <- c(names(df[1:3]), 'pval.lat', 'W', 'U', 'delta')
qval.lat <- p.adjust(lat.100$pval.lat, method='BH') #multiple testing correction
lat.100 <- cbind(lat.100, qval.lat)

lat.100 <- lat.100[which(lat.100$qval.lat <= 0.05),] #keep only significant
```



#SUBSI

Here we apply the same logic to subsistence strategies, this time using the subsi
info inside of the context_pop.txt files. However, the dataset is getting 
restrained to only NCA populations. 
For this, we developed 2 methods 1) NP test and 2) subsi-only specific tests. 

## Method 1: non parametric test
Computing how many windows are being tested (containing only North pops)

```{r}
North <- pops[1:12] 
compteur <- 0 #counts the number of windows containing North-only pops
pas_North <- 0

df <- prop.snp.100

#l circles through the lines of prop_snp
for(l in c(1:nrow(prop.snp.100)))
{
  pops_line <- unlist(strsplit(prop.snp.100$list[l], split= ' '))
  length_pops <- as.numeric(length(pops_line))
  if(length(pops_line)==length(which(pops_line %in% North))){
    compteur <- compteur+1
  }
  else {
    pas_North <- pas_North+1
  }
}
```


```{r NP subsi, warning=FALSE}
#init
df <- prop.snp.100
df <- df[,1:15]
subsi.100 <- data.frame()

#w cycles through the windows of prop_snp
for(w in c(1:dim(df)[1]))
{
  prop_snp <-c(unlist(df[w,4:15]))
  prop_snp <- as.numeric(prop_snp)
  data_lm <- cbind(prop_snp, pop.info[1:12,])
  data_lm <- data_lm[which(data_lm$lat_cat == 'North'),]
  
  #Condition : more than 2 values are different from 0
  if(as.numeric(length(which(data_lm$prop_snp != 0))) >= 2 & table(is.na(data_lm$prop_snp))[[1]] == 12)
  {
    #Farmers vs others----------------------------------------------------------
    
    data_lm$Lifestyle[which(data_lm$Lifestyle == "Hunter-gatherer")] <- c("Herder")
    data_lm$Lifestyle <- as.factor(data_lm$Lifestyle)
    test <- wilcox.test(prop_snp~Lifestyle, data_lm)
    #compute pvalue for latitude
    pval.F <- test$p.value
    # Extract the W statistic
    W.F <- test$statistic
    # Sizes of the groups
    n1 <- sum(data_lm$Lifestyle== "Farmer")
    n2 <- sum(data_lm$lat_cat == "Herder")
    # Calculate U1 and U2
    U1 <- W.F - (n1 * (n1 + 1)) / 2
    U2 <- n1 * n2 - U1
    # Calculate the U statistic (the smaller of U1 and U2)
    U.F <- min(U1, U2)
    size.effect.F <- cliff.delta(prop_snp ~ Lifestyle, data = data_lm)$estimate
    
    data_lm <- cbind(prop_snp, pop.info[1:12,])
  data_lm <- data_lm[which(data_lm$lat_cat == 'North'),]
  
  
  #Herders vs others------------------------------------------------------------
  
  data_lm$Lifestyle[which(data_lm$Lifestyle == "Hunter-gatherer")] <- c("Farmer")
    data_lm$Lifestyle <- as.factor(data_lm$Lifestyle)
    test <- wilcox.test(prop_snp~Lifestyle, data_lm)
    pval.H <- test$p.value
    # Extract the W statistic
    W.H <- test$statistic
    # Sizes of the groups
    n1 <- sum(data_lm$Lifestyle== "Farmer")
    n2 <- sum(data_lm$lat_cat == "Herder")
    # Calculate U1 and U2
    U1 <- W.H - (n1 * (n1 + 1)) / 2
    U2 <- n1 * n2 - U1
    # Calculate the U statistic (the smaller of U1 and U2)
    U.H <- min(U1, U2)
    size.effect.H <- cliff.delta(prop_snp ~ Lifestyle, data = data_lm)$estimate
        
    data_lm <- cbind(prop_snp, pop.info[1:12,])
  data_lm <- data_lm[which(data_lm$lat_cat == 'North'),]
  
  
  #HG vs others----------------------------------------------------------------- 
  data_lm$Lifestyle[which(data_lm$Lifestyle == "Herder")] <- c("Farmer")
  data_lm$Lifestyle <- as.factor(data_lm$Lifestyle)
  test <- wilcox.test(prop_snp~Lifestyle, data_lm)  
  pval.HG <- test$p.value
    # Extract the W statistic
    W.HG <- test$statistic
    # Sizes of the groups
    n1 <- sum(data_lm$Lifestyle== "Farmer")
    n2 <- sum(data_lm$lat_cat == "Hunter-gatherer")
    # Calculate U1 and U2
    U1 <- W.HG - (n1 * (n1 + 1)) / 2
    U2 <- n1 * n2 - U1
    # Calculate the U statistic (the smaller of U1 and U2)
    U.HG <- min(U1, U2)
    size.effect.HG <- cliff.delta(prop_snp ~ Lifestyle, data = data_lm)$estimate
  
  } 
  else 
  {
    pval.F <- NA
    pval.H <- NA
    pval.HG <- NA
    W.F <- NA
    W.H <- NA
    W.HG <- NA
    U.F <- NA
    U.H <- NA
    U.HG <- NA
    size.effect.F <- NA
    size.effect.H <- NA
    size.effect.HG <- NA
  }
  subsi.100 <- rbind(subsi.100, c(df[w,1:3], pval.F, W.F, U.F, size.effect.F,pval.H, W.H, U.H, size.effect.H, pval.HG, W.HG, U.HG, size.effect.HG ))
  names(subsi.100) <- NULL
}
names(subsi.100) <- c(names(df[1:3]), 'pval.F', 'W.F', 'U.F', 'delta.F', 'pval.H', 'W.H', 'U.H', 'delta.H', 'pval.HG', 'W.HG', 'U.HG', 'delta.HG')
#removing NA values
subsi.100 <- subsi.100[which(!is.na(subsi.100$pval.F)),]

qval.F <- p.adjust(subsi.100$pval.F, method='BH') #multiple testing corrections
qval.H <- p.adjust(subsi.100$pval.H, method='BH')
qval.HG <- p.adjust(subsi.100$pval.HG, method='BH')
subsi.100 <- cbind(subsi.100, qval.F, qval.H, qval.HG)

F.wil <- subsi.100[which(subsi.100$qval.F <= 0.05),] #farmer significant
H.wil <- subsi.100[which(subsi.100$qval.H <= 0.05),] #herder significant
HG.wil <- subsi.100[which(subsi.100$qval.HG <= 0.05),] #HG significant 
```

## Method 2: finding subsi-specific windows 
This second method is designed to find the windows that are present only in the 
pops sharing a subsi and absent in the other. We can adjust the level of string-
ency using "pops.pick.H". In the paper, we have used 4/6 for herders, 3/3 for HG
and Farmers. 

```{r finding subsi specific windows}
set.seed(666)

#HERDERS -----------------------------------------------------------------------

pops.pick.H <- 4 #adjust stringency 

#picking x number of herders out of 6
H <- c("AKI","AKZ","BOU","KAZ","KIB","TLG") #name of subset of pops
comb.H <- combinations(length(H),pops.pick.H,H)
comb.H <- as.data.frame(comb.H)
H.undersample <- c()

#l counts the number of H pop combinations possible 
for(l in c(1:nrow(comb.H)))
{
  inter <- unlist(c(comb.H[l,]))
  inter <- inter[order(inter)] #alphabetical 
  H.undersample <- c(H.undersample, paste(inter, collapse=","))
}

any <- c()
spe <- c()
H.spe.win <- data.frame()
for(c in c(1:length(H.undersample)))
{
  any <- c(any, length(which(apply(sapply(H.undersample[c], grepl, intersect.100$list),1, all) == T)))
  spe.indexes <- grep(paste(H.undersample[c], collapse=","), intersect.100$list)
  spe <- c(spe, length(which(intersect.100$num[spe.indexes] == pops.pick.H)))
  spe.indexes <- spe.indexes[which(intersect.100$num[spe.indexes] == pops.pick.H)]
  H.spe.win <- rbind(H.spe.win, intersect.100[spe.indexes,])
}
#all containing
print(paste0(pops.pick.H,"/6 = ", sum(any) , " all"))
#specific
print(paste0(pops.pick.H,"/6 = ", sum(spe) , " specific"))



#HG ----------------------------------------------------------------
pops.pick.F.HG <- 2
#picking 2 F or HG out of 3

HG <- c("HKS", "SHO", "TUB")

comb.HG <- combinations(length(HG),pops.pick.F.HG,HG)
comb.HG <- as.data.frame(comb.HG)
HG.undersample <- c()


for(l in c(1:nrow(comb.HG)))
{
  inter <- unlist(c(comb.HG[l,]))
  inter <- inter[order(inter)]
  HG.undersample <- c(HG.undersample, paste(inter, collapse=","))
}

any <- c()
spe <- c()
F.spe.win <- data.frame()
for(c in c(1:length(HG.undersample)))
{
  any <- c(any, length(which(apply(sapply(HG.undersample[c], grepl, intersect.100$list),1, all) == T)))
  spe.indexes <- grep(paste(HG.undersample[c], collapse=","), intersect.100$list)
  spe <- c(spe, length(which(intersect.100$num[spe.indexes] == pops.pick.F.HG)))
  spe.indexes <- spe.indexes[which(intersect.100$num[spe.indexes] == pops.pick.F.HG)]
  F.spe.win <- rbind(F.spe.win, intersect.100[spe.indexes,])
}
#all containing
print(paste0("HG: ", pops.pick.F.HG,"/3 = ", sum(any) , " all"))
#specific
print(paste0("HG: ", pops.pick.F.HG,"/3 = ", sum(spe) , " specific"))


#FARMERS -----------------------------------------------------------------------
pops.pick.F.HG <- 3
Farm <- c("TAB", "TUR", "UZB")
comb.F <- combinations(length(Farm),pops.pick.F.HG,Farm)
comb.F <- as.data.frame(comb.F)
F.undersample <- c()

for(l in c(1:nrow(comb.F)))
{
  inter <- unlist(c(comb.F[l,]))
  inter <- inter[order(inter)]
  F.undersample <- c(F.undersample, paste(inter, collapse=","))
}

any <- c()
spe <- c()
F.spe.win <- data.frame()
for(c in c(1:length(F.undersample)))
{
  any <- c(any, length(which(apply(sapply(F.undersample[c], grepl, intersect.100$list),1, all) == T)))
  spe.indexes <- grep(paste(F.undersample[c], collapse=","), intersect.100$list)
  spe <- c(spe, length(which(intersect.100$num[spe.indexes] == pops.pick.F.HG)))
  spe.indexes <- spe.indexes[which(intersect.100$num[spe.indexes] == pops.pick.F.HG)]
  F.spe.win <- rbind(F.spe.win, intersect.100[spe.indexes,])
}
#all containing
print(paste0("F: ", pops.pick.F.HG,"/3 = ", sum(any) , " all"))
#specific
print(paste0("F: ", pops.pick.F.HG,"/3 = ", sum(spe) , " specific"))

```

##bootstrap for Method 2

For this boostrap, I enter the values of the number of windows found in for 6/6
Herder, 5/6 Herder, 4/6 Herder etc as input. 
```{r bootstrap subsi}
set.seed(666)


#HERDERS------------------------------------------------------------------------
results.H.spe <- c(0,0,9,18,98) #number of windows when doing the "real" test
results.H.all <- c(0,3,32,161,643)
r <- 1

for(n in c(6:2))
{
  pops.pick.H <- n
  print(pops.pick.H)
  
  #make a list of forbidden combinations because they have real association
  H.boot <- c("AKI","AKZ","BOU","KAZ","KIB","TLG")
  forbidden.comb <- combinations(6,pops.pick.H,H.boot) #bootstrap for 6/6
  forbidden.comb <- as.data.frame(forbidden.comb)
  forbidden <- c()
  for(l in c(1:nrow(forbidden.comb)))
  {
    inter <- unlist(c(forbidden.comb[l,]))
    inter <- inter[order(inter)]
    forbidden <- c(forbidden, paste(inter, collapse=","))
  }
  
  
  any.H <- c()
  spe.H <- c()
  reshuffle.H <- 0
  
  for(b in c(1:10000)){
    #herder=6
    H.boot <- sample(pops, pops.pick.H)
    H.boot <- H.boot[order(H.boot)]
    #as long as the random combination is "forbidden", reshuffle
    while(paste(H.boot, collapse=",") %in% forbidden) 
    {
      reshuffle.H <- reshuffle.H+1
      H.boot <- sample(pops, pops.pick.H)
      H.boot <- H.boot[order(H.boot)]
    }
    #find any list containing the names in HG.boot 
    any.H <- c(any.H, length(which(apply(sapply(H.boot, grepl, intersect.100$list),1, all) == T)))
    #find any list containing only the names in HG.boot 
    spe.H <- c(spe.H, length(grep(paste(H.boot, collapse=","), intersect.100$list)))
  }
  #all containing
  print(paste0(pops.pick.H,"/6 = ", mean(any.H) , " all. Details: max=", max(any.H), " min=", min(any.H)))
  print(paste0("sup:", length(which(any.H >= results.H.all[r]))))
  #specific
  print(paste0(pops.pick.H,"/6 = ", mean(spe.H) , " specific. Details: max=", max(spe.H), " min=", min(spe.H)))
  print(paste0("sup:", length(which(spe.H >= results.H.spe[r]))))
  print(paste0("reshuffle = ", reshuffle.H))
  r <- r+1
}

#HG-----------------------------------------------------------------------------
set.seed(666)
results.HG.all <- c(1,40)
results.HG.spe <- c(1,21)
r <- 1
for(n in c(3:2))
{
  pops.pick.F.HG <- n
  print(pops.pick.F.HG)
  #picking 2 F or HG out of 3
  HG.boot <- c("HKS", "SHO", "TUB")
  forbidden.comb.HG <- combinations(3,pops.pick.F.HG,HG.boot)
  forbidden.comb.HG <- as.data.frame(forbidden.comb.HG)
  forbidden.HG <- c()
  for(l in c(1:nrow(forbidden.comb.HG)))
  {
    inter <- unlist(c(forbidden.comb.HG[l,]))
    inter <- inter[order(inter)]
    forbidden.HG <- c(forbidden.HG, paste(inter, collapse=","))
  }
  
  any.F.HG <- c()
  spe.F.HG <- c()
  reshuffle.HG.F <- 0
  for(b in c(1:10000)){
    #HG or Farmers = 3
    HG.F.boot <- sample(pops, pops.pick.F.HG)
    HG.F.boot <- HG.F.boot[order(HG.F.boot)]  #alphabetical to match order in intersect.100$list
    #resampling to avoid these combinations
    while(paste(HG.F.boot, collapse=",") %in% forbidden.HG)
    {
      reshuffle.HG.F <- reshuffle.HG.F+1
      HG.F.boot <- sample(pops, pops.pick.F.HG)
      HG.F.boot <- HG.F.boot[order(HG.F.boot)]
    }
    #find any list containing the names in HG.boot 
    any.F.HG <- c(any.F.HG, length(which(apply(sapply(HG.F.boot, grepl, intersect.100$list),1, all) == T)))
    #find any list containing only the names in HG.boot 
    spe.F.HG <- c(spe.F.HG, length(grep(paste(HG.F.boot, collapse=","), intersect.100$list)))
  }
  #all containing
  print(paste0("HG: ", pops.pick.F.HG,"/3 = ", mean(any.F.HG) , " all"))
  print(paste0("sup:", length(which(any.F.HG >= results.HG.all[r]))))
  #specific
  print(paste0("HG: ", pops.pick.F.HG,"/3 = ", mean(spe.F.HG) , " specific. Max=", max(spe.F.HG), " min = ", min(spe.F.HG)))
  print(paste0("sup:", length(which(spe.F.HG >= results.HG.spe[r]))))
  print(paste0("reshuffle = ", reshuffle.HG.F))
  r <- r+1
}

#F------------------------------------------------------------------------------
set.seed(666)
results.HG.all <- c(27,204)
results.HG.spe <- c(3,40)
r <- 1
for(n in c(3:2))
{
  pops.pick.F.HG <- n
  print(pops.pick.F.HG)
  #picking 2 F or HG out of 3
  HG.boot <- c("TAB", "TUR", "UZB")
  forbidden.comb.HG <- combinations(3,pops.pick.F.HG,HG.boot)
  forbidden.comb.HG <- as.data.frame(forbidden.comb.HG)
  forbidden.HG <- c()
  for(l in c(1:nrow(forbidden.comb.HG)))
  {
    inter <- unlist(c(forbidden.comb.HG[l,]))
    inter <- inter[order(inter)]
    forbidden.HG <- c(forbidden.HG, paste(inter, collapse=","))
  }
  
  any.F.HG <- c()
  spe.F.HG <- c()
  reshuffle.HG.F <- 0
  for(b in c(1:10000)){
    #HG or Farmers = 3
    HG.F.boot <- sample(pops, pops.pick.F.HG)
    HG.F.boot <- HG.F.boot[order(HG.F.boot)]  #alphabetical to match order in intersect.100$list
    #resampling to avoid these combinations
    while(paste(HG.F.boot, collapse=",") %in% forbidden.HG)
    {
      reshuffle.HG.F <- reshuffle.HG.F+1
      HG.F.boot <- sample(pops, pops.pick.F.HG)
      HG.F.boot <- HG.F.boot[order(HG.F.boot)]
    }
    #find any list containing the names in HG.boot 
    any.F.HG <- c(any.F.HG, length(which(apply(sapply(HG.F.boot, grepl, intersect.100$list),1, all) == T)))
    #find any list containing only the names in HG.boot 
    spe.F.HG <- c(spe.F.HG, length(grep(paste(HG.F.boot, collapse=","), intersect.100$list)))
  }
  #all containing
  print(paste0("HG: ", pops.pick.F.HG,"/3 = ", mean(any.F.HG) , " all"))
  print(paste0("sup:", length(which(any.F.HG >= results.HG.all[r]))))
  #specific
  print(paste0("HG: ", pops.pick.F.HG,"/3 = ", mean(spe.F.HG) , " specific. Max=", max(spe.F.HG), " min = ", min(spe.F.HG)))
  print(paste0("sup:", length(which(spe.F.HG >= results.HG.spe[r]))))
  print(paste0("reshuffle = ", reshuffle.HG.F))
  r <- r+1
}
```
